Main Analysis for Biracial Grade Retention Project

### 2022-09-04 analysis/analysis_main.Rmd
Refactor analysis.Rmd into three separate Rmd files

---
title: "Analysis for Biracial Grade Retention Project"
output: 
  rmdformats::robobook:
    highlight: tango
    toc_depth: 2
    fig_height: 6
    fig_width: 9
    code_folding: hide
    lightbox: TRUE
---

@@

# Grade retention summary

OK, first lets just confirm that the grade retention variable is working as expected on our sample. I start by looking at the number of kids by age and current grade:

```{r tab-grade-by-age}
acs %>% 
  tabyl(age, current_grade) %>%
  mutate_if(is.numeric, format, big.mark=",") %>%
  kbl(caption="Cross-tabulation of age and current grade of student",
      align=c("l",rep("r",13))) %>%
  kable_paper() %>%
  column_spec(2, background=c(rep("#D3D3D3",2), rep("white", 14))) %>%
  column_spec(3, background=c("white", 
                              rep("#D3D3D3",2), 
                              rep("white", 13))) %>%
  column_spec(4, background=c(rep("white", 2), 
                              rep("#D3D3D3",2), 
                              rep("white", 12))) %>%
  column_spec(5, background=c(rep("white", 3), 
                              rep("#D3D3D3",2), 
                              rep("white", 11))) %>%
  column_spec(6, background=c(rep("white", 4), 
                              rep("#D3D3D3",2), 
                              rep("white", 10))) %>%
  column_spec(7, background=c(rep("white", 5), 
                              rep("#D3D3D3",2), 
                              rep("white", 9))) %>%
  column_spec(8, background=c(rep("white", 6), 
                              rep("#D3D3D3",2), 
                              rep("white", 8))) %>%
  column_spec(9, background=c(rep("white", 7), 
                              rep("#D3D3D3",2), 
                              rep("white", 7))) %>%
  column_spec(10, background=c(rep("white", 8), 
                               rep("#D3D3D3",2), 
                               rep("white", 6))) %>%
  column_spec(11, background=c(rep("white", 9), 
                               rep("#D3D3D3",2), 
                               rep("white", 5))) %>%
  column_spec(12, background=c(rep("white", 10), 
                               rep("#D3D3D3",2), 
                               rep("white", 4))) %>%
  column_spec(13, background=c(rep("white", 11), 
                               rep("#D3D3D3",2), 
                               rep("white", 3))) %>%
  column_spec(14, background=c(rep("white", 12), 
                               rep("#D3D3D3",2), 
                               rep("white", 2)))
```

Ok, this looks more or less like what I would expect. We observe two modal ages for each grade. Some of the more extreme age values for older grades seem a little strange given that these are only supposed to be currently enrolled students. 20-year old sixth graders? I mean there are relatively few of these cases but still seems a bit odd. Its also interesting how this really starts expanding past 5th - did the Census Bureau "correct" values below 6th or something?

Now lets calculate the percentage of cases TRUE on the dependent variable across this same table, just to confirm its all working correctly.

```{r check-retention-code, fig.cap="The proportion of students measured at below grade level by age and current grade. This graph is used to ensure that the dependent variable is calculated correctly."}
acs %>%
  group_by(age, current_grade) %>%
  summarize(below=mean(below_exp_grade)) %>%
  ggplot(aes(x=current_grade, y=age, fill=below))+
  geom_tile(color="grey20")+
  geom_text(aes(label=below))+
  theme_bw()+
  scale_fill_viridis_c()+
  scale_y_continuous(breaks=5:20, minor_breaks = NULL)+
  labs(x="current grade", y="age", fill="proportion below grade level")
```

Ok, that looks good.

What do the grade retention probabilities look like by grade:

```{r retention-by-grade, fig.cap="Percent of students clearly behind grade level by current grade. Blue lines fit a smoothed curve to the data."}
acs %>%
  group_by(current_grade) %>%
  summarize(below_grade=mean(below_exp_grade),
            n=n(),
            se=sqrt((below_grade*(1-below_grade))/n)) %>%
  ggplot(aes(x=current_grade, y=below_grade, group=1,
             ymin=below_grade-1.96*se, ymax=below_grade+1.96*se))+
  geom_smooth(se=FALSE)+
  geom_line(linetype=2, color="grey30")+
  geom_point()+
  geom_linerange()+
  scale_y_continuous(labels=scales::percent)+
  labs(x="current grade", y="percent clearly behind grade level for age")+
  theme_bw()
```

I guess those numbers look pretty reasonable. Its hard to benchmark it off anything. Warren et. al. estimate annual retention rates by grade but that is different than what I am measuring here because this is cumulative (but can also reflect students being caught up). In general, these must be underestimates because they do not capture kids who were held back but are still within the modal age.

What about by year?

```{r retention-by-year, fig.cap="Trends over time in the percent of students clearly behind grade level. Blue lines fit a smoothed curve to the data."}
acs %>%
  group_by(year) %>%
  summarize(below_grade=mean(below_exp_grade)) %>%
  ggplot(aes(x=year, y=below_grade))+
  geom_line()+
  geom_point()+
  geom_smooth(se=FALSE)+
  scale_x_continuous(breaks=c(2010,2014,2018))+
  scale_y_continuous(labels=scales::percent)+
  labs(x="current grade", y="percent clearly behind grade level for age")+
  theme_bw()
```

Overall, there has been some decline over time in the prevalence of grade retention.

Lets look at this by grade and year

```{r retention-by-year-grade, fig.cap="Trends over time and grade in the percent of students clearly behind grade level. Lines are calculated by smoothing."}
acs %>%
  group_by(current_grade, year) %>%
  summarize(below_grade=mean(below_exp_grade)) %>%
  ggplot(aes(x=year, y=below_grade, group=current_grade, color=current_grade))+
  #geom_line()+
  geom_smooth(se=FALSE)+
  geom_point()+
  scale_x_continuous(breaks=c(2010,2014,2018))+
  scale_y_continuous(labels=scales::percent)+
  labs(x="current grade", y="percent clearly behind grade level for age")+
  theme_bw()
```

The evidence here suggests that the decline is driven by the higher grades where grade retention was higher overall.

@@

## Functional form of year-grade interactions

Before running the full models, I want to test out how I can parsimoniously address differences in the time trends by grade.

```{r models-year-grade}
model_add_cat <- glm(below_exp_grade~as.factor(year)+current_grade+state, 
                     data=acs, family=binomial)
model_add_cont <- glm(below_exp_grade~I(year-2010)+current_grade+state, 
                     data=acs, family=binomial)
model_inter_cont <- glm(below_exp_grade~I(year-2010)*current_grade+state, 
                     data=acs, family=binomial)
model_inter_cat <- glm(below_exp_grade~as.factor(year)*current_grade+state, 
                     data=acs, family=binomial)
model_inter_state <- glm(below_exp_grade~I(year-2010)*current_grade+
                            I(year-2010)*state, 
                     data=acs, family=binomial)

sapply(list("categorical, additive"=model_add_cat,
            "categorical, interaction"=model_inter_cat, 
            "continuous, additive"=model_add_cont, 
            "continuous, interaction"=model_inter_cont,
            "continous, interaction + state*year continuous interaction"=model_inter_state),
       BIC) %>%
    enframe("model","BIC") %>%
  kbl() %>%
  kable_styling(full_width=FALSE)

# free some memory
rm(model_add_cat, model_add_cont, model_inter_cat, model_inter_cont,
   model_inter_state)
```

The preferred model by BIC is clearly the model that fits year as a continuous measure and interacts it with current grade, so that is what I will do.

@@

# Differences in Family Resources

## Full tables of SES and cultural resources

First, lets compare means of SES and cultural resources across all racial groups.

```{r desc-stats-ses}
tbl_summary <- acs_design %>%
  group_by(race) %>%
  summarize(fam_inc=survey_mean(family_income, vartype = NULL))

tbl_summary <- acs_design %>%
  group_by(race) %>%
  summarize(fam_inc=round(survey_mean(family_income, vartype = NULL), 0),
            own_home=round(survey_mean(own_home, vartype = NULL), 3)*100,
            par_mar=round(survey_mean(parents_married, vartype = NULL), 3)*100,
            f_lhs=round(survey_mean(degree_father=="LHS", vartype = NULL), 
                        3)*100,
            f_hs=round(survey_mean(degree_father=="HS", vartype = NULL), 3)*100,
            f_aa=round(survey_mean(degree_father=="AA", vartype = NULL), 3)*100,
            f_ba=round(survey_mean(degree_father=="BA" | degree_father=="G", 
                                   vartype = NULL), 3)*100,
            m_lhs=round(survey_mean(degree_mother=="LHS", vartype = NULL), 3)*100,
            m_hs=round(survey_mean(degree_mother=="HS", vartype = NULL), 3)*100,
            m_aa=round(survey_mean(degree_mother=="AA", vartype = NULL), 3)*100,
            m_ba=round(survey_mean(degree_mother=="BA" | degree_mother=="G", 
                                   vartype = NULL), 3)*100)

tbl_summary %>%
  mutate_if(is.numeric, format, big.mark=",") %>%
  kbl(caption="SES resources by race. Shading indicates biracial group.",
      align = c("l",rep("r", ncol(tbl_summary)-1)),
      col.names=c("race","mean family income","% own home",
                  "% parents married",
                  "% father, no HS diploma","% father, HS diploma",
                  "% father, AA degree",  "% father, BA+ degree",
                  "% mother, no HS diploma","% mother, HS diploma", 
                  "%  mother, AA degree", "% mother, BA+ degree")) %>%
  kable_paper() %>%
  row_spec(index_biracial, background = "#D3D3D3")
```

There is a lot going on here, but I can summarize a few things:

- Most multiracial groups are more like the more privileged constituent group than the less privileged constituent group across all of these characteristics. For example, White/Latino respondents have family incomes of \$130,981 compared to \$138,489 for Whites and \$64,920 for Latinos. Similarly, 9.6% and 11.6% of Latino fathers and mothers, respectively, have a BA degree, compared to 43.8% and 47.6% for Whites, and 38.6% and 41.5% for White/Latinos. Also White/Latino parents own their home in 86.7% of cases which is similar to the white value of 85.5% and much higher than the 54.3% for Latinos.
- Black/Whites in contrast, are much more like Blacks in terms of their resources and both groups have fewer resources than Whites.
- White/Asians are the most privileged group of all, even more so than their two constituent groups. Their family income of \$195,028 is off the charts! but they also have parents with very high educational attainment.
- Most part-Indigenous groups also buck the modal trend and have outcomes that are about halfway between their two constituent groups. The one part-Indigenous group that does not follow that pattern is the Indigenous/Latino group, which tend to have resources slightly better than both of the monoracial groups (although still very low compared to everyone else).

Now lets look at the cultural characteristics. One thing to keep in mind here is that having foreign-born parents actually *lowers* your risk of grade retention, whereas being foreign-born yourself increases it.

```{r desc-stats-cult}
tbl_summary <- acs_design %>%
  group_by(race) %>%
  summarize(f_born=round(survey_mean(foreign_born, vartype = NULL), 3)*100,
            father_fb=round(survey_mean(foreign_born_father, 
                                        vartype = NULL), 3)*100,
            mother_fb=round(survey_mean(foreign_born_mother, 
                                        vartype = NULL), 3)*100,
            spk_eng_no=round(survey_mean(spk_eng=="No", vartype = NULL), 3)*100,
            spk_eng_some=round(survey_mean(spk_eng=="Somewhat", 
                                           vartype = NULL), 3)*100,
            spk_eng_yes=round(survey_mean(spk_eng=="Yes", vartype = NULL), 3)*100,
            f_spk_eng_no=round(survey_mean(spk_eng_father=="No", 
                                           vartype = NULL), 3)*100,
            f_spk_eng_some=round(survey_mean(spk_eng_father=="Somewhat", 
                                             vartype = NULL), 3)*100,
            f_spk_eng_yes=round(survey_mean(spk_eng_father=="Yes", 
                                            vartype = NULL), 3)*100,
            m_spk_eng_no=round(survey_mean(spk_eng_mother=="No", 
                                           vartype = NULL), 3)*100,
            m_spk_eng_some=round(survey_mean(spk_eng_mother=="Somewhat", 
                                             vartype = NULL), 3)*100,
            m_spk_eng_yes=round(survey_mean(spk_eng_mother=="Yes", 
                                            vartype = NULL), 3)*100)

tbl_summary %>%
  mutate_if(is.numeric, format, big.mark=",") %>%
  kbl(caption="Cultural resources by race. Shading indicates biracial group.",
      align = c("l",rep("r", ncol(tbl_summary)-1)),
      col.names=c("race",
                  "% foreign-born","% father foreign-born",
                  "% mother foreign-born",
                  "% speak no English", "% speak some English", 
                  "% speak English",
                  "% father speak no English", "% father speak some English", 
                  "% father speak English",
                  "% mother speak no English", "% mother speak some English",
                  "% mother speak English")) %>%
  kable_paper() %>%
  row_spec(index_biracial, background = "#D3D3D3")
```
* In general, the most notable thing here is that in terms of English proficiency (both respondent and parents), all of the biracial groups involving Latino and Asian have almost universal proficiency (>99%) and more like the non-Latino and non-Asian groups. This is most notable for Latinos where only about two-thirds of fathers and mothers of monoracial Latinos have full English proficiency but all of the part Latino multiracial groups have parents that are 99% or more fluent. So, there is heavy selection on English proficiency into interracial unions, which is not surpising.
* We also see that part-Latino and part-Asian respondents are much more likely to be native born and to have native-born parents than Latino and Asian respondents. The effects are large although not quite as huge as for language. However, because foreign-born parents actually *reduce* the effect of grade retention, the ultimate effect of this selection is more complex.

## Counterfactual models by resource differentials

One way to boil down the complexity of these tables into something easier to understand would be to calculate average predicted probability for each racial group, while holding constant all non-resource variables *including race*. These probabilities provide a counterfactual estimate of what the average probability would be for each group if they only differed on the resource variables, allowing me to isolate how different distributions of resources across racial groups affect the risk of grade retention.

To do this, the following variables need to be set at some constant value to be held constant:

- current grade
- state
- year
- metro status
- gender
- race

One could just plug in some static values (e.g. make everyone White, from Minnesota, in 6th grade, etc.), but this feels pretty ad-hoc. I think the best approach is to set all of these values at their mean. Since most of these variables are categorical variables, this means treating them as 0/1 dummies and then calculating the proportion as the mean. Unfortunately, this approach won't work with the models built above because they made use of factor variables. So, to run these models, I first need to create actual 0/1 dummies for the categorical variables listed above (current grade, state, metro status, gender, race). That is what the code below does.

```{r create-dummied-data}
acs_dummied <- acs %>%
  # first we need to unorder current grade for dummy_cols to work
  mutate(current_grade=factor(current_grade, levels=levels(current_grade), 
                              ordered = FALSE)) %>%

  dummy_cols(remove_first_dummy = TRUE) %>%
  select(cluster, strata, perwt, below_exp_grade, year, family_income, 
         parents_married, own_home, degree_mother, degree_father,
         foreign_born, foreign_born_mother, foreign_born_father, 
         spk_eng, spk_eng_father, spk_eng_mother,
         starts_with(c("state_","current_grade_","race","sex_","metro_"))) %>%
  mutate(year=year-2010) %>%
  clean_names()
```

Now that I have new "dummied up" dataset, I can rereun the full model using the dummy variables which means a very, very long formula.

```{r model-dummied}
acs_dummied_design <- svydesign(ids=~cluster, weights=~perwt, data=acs_dummied)

model_full <- svyglm(below_exp_grade~
                       year*(current_grade_1st+current_grade_2nd+current_grade_3rd+
                               current_grade_4th+current_grade_5th+
                               current_grade_6th+current_grade_7th+
                               current_grade_8th+current_grade_9th+
                               current_grade_10th+current_grade_11th+
                               current_grade_12th)+
                       race_black+race_indigenous+race_asian+race_latino+
                       race_black_white+race_black_indigenous+race_black_latino+
                       race_black_asian+race_white_indigenous+race_white_latino+
                       race_white_asian+race_indigenous_latino+
                       race_indigenous_asian+race_latino_asian+
                       sex_female+
                       metro_non_metro+metro_central_city+metro_metro_non_central+
                       metro_metro_indeterminable+
                       state_alaska+state_arizona+state_arkansas+state_california+
                       state_colorado+state_connecticut+state_dc+state_delaware+
                       state_florida+state_georgia+state_hawaii+state_idaho+
                       state_illinois+state_indiana+state_iowa+state_kansas+
                       state_kentucky+state_louisiana+state_maine+state_maryland+
                       state_massachusetts+state_michigan+state_minnesota+
                       state_mississippi+state_missouri+state_montana+
                       state_nebraska+state_nevada+state_new_hampshire+
                       state_new_jersey+state_new_mexico+state_new_york+
                       state_north_carolina+state_north_dakota+state_ohio+
                       state_oklahoma+state_oregon+state_pennsylvania+
                       state_rhode_island+state_south_carolina+state_south_dakota+
                       state_tennessee+state_texas+state_utah+state_vermont+
                       state_virginia+state_washington+state_west_virginia+
                       state_wisconsin+state_wyoming+
                       +sqrt(family_income/10000)+parents_married+own_home+
                       degree_mother+degree_father+
                       spk_eng+spk_eng_mother+spk_eng_father+
                       foreign_born_mother+foreign_born_father+foreign_born, 
                     design=acs_dummied_design, family=binomial)
```

I now have the model in the correct form. The next step is to create the counterfactual dataset by setting all of the non-resource variables to their mean. Thus, in this dataset, the only differences across racial groups are coming from their resource differences.

```{r create-cf-df}
# Create the counterfactual dataset where non-resource variables are held
# at their mean. What needs to be set at tis mean?
# - current_grade
# - sex
# - race
# - state
# - metro
# - year
means <- acs_dummied %>%
  select(year, starts_with(c("state_","current_grade_","race_","sex_",
                             "metro_"))) %>%
  lapply(function(x) {rep(mean(x), length(x))}) %>%
  bind_rows()

cf_df <- acs_dummied %>%
  select(cluster, strata, perwt, below_exp_grade, race,family_income, 
         parents_married, own_home, degree_mother, degree_father, foreign_born, 
         foreign_born_mother, foreign_born_father, spk_eng, spk_eng_father, 
         spk_eng_mother) %>%
  bind_cols(means)
```

```{r test-marg, eval=FALSE}
# This code chunk just tests to make sure I get the same value by hand as I get 
# when I do it through the marg command.
temp <- subset(cf_df, race=="Black/White")
lor <- predict(model_full, newdata=temp)
weighted.mean(exp(lor)/(1+exp(lor)), w=temp$perwt)
marg(model_full, "year", data=temp, weights=temp$perwt)
# It works
```

I can then feed subsets (by race) of this counterfactual dataset into the `marg` command to get the predicted probability of grade retention for each racial group as I do below.

```{r create-cf-marg}
coefs_cf <- lapply(unique(acs$race), function(x) {
  temp <- subset(cf_df, race==x)
  means_margin <- marg(model_full, "year", data=temp, weights=temp$perwt)[[1]]
  return(tibble(term=x, estimate=means_margin$Margin, 
                se=means_margin$Standard.Error))
}) %>%
  bind_rows()

coef_table_cf <- create_coef_table(coefs_cf)
```

And now I can plot up the results.

```{r counterfactual-ses, warning=FALSE, fig.cap="Counterfactual probability of being behind grade for biracial respondents in comparison to their monoracial comparison groups if the only differences between groups were resources (household income, parental education, home ownership, parent and child foreign-born status, parent and child English proficiency, and whether parents are married). 83.4% confidence bands shown around each estimate. Non-overlap in confidence bands roughly indicates statistically significant difference at the 5% level."}
ggplot(coef_table_cf, aes(x=term, y=estimate, color=multiracial,
                          ymin=estimate-1.386*se, ymax=estimate+1.386*se))+
  geom_rect(aes(xmax=Inf, xmin=-Inf, ymin=estimate_mid-1.386*se_mid,
                ymax=estimate_mid+1.386*se_mid, fill=multiracial), 
            color=NA, alpha=0.2)+
  geom_linerange(data=subset(coef_table_cf, multiracial))+
  geom_point()+
  coord_flip()+
  facet_wrap(~mrace, scales="free_y")+
  theme_bw()+
  theme(legend.position = "none", panel.grid = element_blank())+
  scale_color_manual(values=c("grey40","red"))+
  scale_fill_manual(values=c("grey40","red"))+
  scale_linetype_manual(values=c(3,1))+
  scale_y_continuous(labels = scales::percent)+
  labs(x=NULL,
       y="average adjusted predicted probability of being behind expected grade")

```

I summarize the results here:

|                   | Lower than both | Same as lower | Between | Same as higher | Higher than both |
|-------------------|:---------------:|:-------------:|:-------:|:--------------:|:----------------:|
| White/Asian       |       X         |               |         |                |                  |
| Black/Asian       |                 |  X  (Asian)   |         |                |                  |
| Black/Latino      |                 |  X  (Black)   |         |                |                  |
| White/Latino      |                 |  X  (White)   |         |                |                  |
| Indigenous/Latino |                 |X  (Indigenous)|         |                |                  |
| Latino/Asian      |                 |  X (Asian)    |         |                |                  |
| White/Indigenous  |                 |               |    X    |                |                  |
| Indigenous/Asian  |                 |               |    X    |                |                  |
| Black/Indigenous  |                 |               |    X    |                |                  |
| Black/White       |                 |               |         |   X (Black)    |                  |

In half the cases, the biracial group would have outcomes similar to the group with better resources suggesting that the parents in these matches are being positively selected. The White/Asian case is even more dramatic with a resource advantage over both single race groups. You can see this particularly for the White/Asian case in terms of family income where they have the highest of *any* group by a considerable amount. White/Asians are on average a very privileged group.

Three groups that have the "expected" in-between status based on resources all share one thing in common - they are part-Indigenous. That is not true of the Indigenous/Latino case, but is of all others. It suggests less selectivity in crossing this boundary.

Finally, the Black/White case stands out in that Black/White kids have the same risk of being held back based on their resources as do Black kids. Thus, they are not getting a resource boost nor does it seem like they are particularly positively selected.

So the effects of controlling for these resources will be complex and different across groups. In general for those that are the same as lower or lower than both, we see that it pulls their relative probability up somewhat. For the Black/White case, it pulls their probability down on the other hand. For the "between" groups its more complex and doesn't seem to do much in total.

### Senstivity tests separating SES and cultural differences

Lets do the same procedure but I want to isolate the role of SES vs cultural (foreign-born and language).

For ease of use, I am just going to plug in values here for the other resources rather than dummy everything up.

```{r cf-ses-isolate, warning=FALSE, fig.cap="Counterfactual probability of being behind grade for biracial respondents in comparison to their monoracial comparison groups if the only differences between groups were SES resources (household income, parental education, home ownership, and whether parents are married). 83.4% confidence bands shown around each estimate. Non-overlap in confidence bands roughly indicates statistically significant difference at the 5% level."}
cf_df_ses <- cf_df
cf_df_ses$foreign_born <- FALSE
cf_df_ses$foreign_born_father <- FALSE
cf_df_ses$foreign_born_mother <- FALSE
cf_df_ses$spk_eng <- "Yes"
cf_df_ses$spk_eng_father <- "Yes"
cf_df_ses$spk_eng_mother <- "Yes"

coefs_cf_ses <- lapply(unique(acs$race), function(x) {
  temp <- subset(cf_df_ses, race==x)
  means_margin <- marg(model_full, "year", data=temp, weights=temp$perwt)[[1]]
  return(tibble(term=x, estimate=means_margin$Margin, 
                se=means_margin$Standard.Error))
}) %>%
  bind_rows()

coef_table_cf_ses <- create_coef_table(coefs_cf_ses)

ggplot(coef_table_cf_ses, aes(x=term, y=estimate, color=multiracial,
                          ymin=estimate-1.386*se, ymax=estimate+1.386*se))+
  geom_rect(aes(xmax=Inf, xmin=-Inf, ymin=estimate_mid-1.386*se_mid,
                ymax=estimate_mid+1.386*se_mid, fill=multiracial), 
            color=NA, alpha=0.2)+
  geom_linerange(data=subset(coef_table_cf_ses, multiracial))+
  geom_point()+
  coord_flip()+
  facet_wrap(~mrace, scales="free_y")+
  theme_bw()+
  theme(legend.position = "none", panel.grid = element_blank())+
  scale_color_manual(values=c("grey40","red"))+
  scale_fill_manual(values=c("grey40","red"))+
  scale_linetype_manual(values=c(3,1))+
  scale_y_continuous(labels = scales::percent)+
  labs(x=NULL,
       y="average adjusted predicted probability of being behind expected grade")
```

```{r cf-culture-isolate, warning=FALSE, fig.cap="Counterfactual probability of being behind grade for biracial respondents in comparison to their monoracial comparison groups if the only differences between groups were cultural resources (foreign-born status and English proficiency). 83.4% confidence bands shown around each estimate. Non-overlap in confidence bands roughly indicates statistically significant difference at the 5% level."}
cf_df_culture <- cf_df
cf_df_culture$family_income <- mean(cf_df_culture$family_income)
cf_df_culture$own_home <- TRUE
cf_df_culture$degree_father <- "BA"
cf_df_culture$degree_mother <- "BA"
cf_df_culture$parents_married <- TRUE

coefs_cf_culture <- lapply(unique(acs$race), function(x) {
  temp <- subset(cf_df_culture, race==x)
  means_margin <- marg(model_full, "year", data=temp, weights=temp$perwt)[[1]]
  return(tibble(term=x, estimate=means_margin$Margin, 
                se=means_margin$Standard.Error))
}) %>%
  bind_rows()

coef_table_cf_culture <- create_coef_table(coefs_cf_culture)

ggplot(coef_table_cf_culture, aes(x=term, y=estimate, color=multiracial,
                          ymin=estimate-1.386*se, ymax=estimate+1.386*se))+
  geom_rect(aes(xmax=Inf, xmin=-Inf, ymin=estimate_mid-1.386*se_mid,
                ymax=estimate_mid+1.386*se_mid, fill=multiracial), 
            color=NA, alpha=0.2)+
  geom_linerange(data=subset(coef_table_cf_culture, multiracial))+
  geom_point()+
  coord_flip()+
  facet_wrap(~mrace, scales="free_y")+
  theme_bw()+
  theme(legend.position = "none", panel.grid = element_blank())+
  scale_color_manual(values=c("grey40","red"))+
  scale_fill_manual(values=c("grey40","red"))+
  scale_linetype_manual(values=c(3,1))+
  scale_y_continuous(labels = scales::percent)+
  labs(x=NULL,
       y="average adjusted predicted probability of being behind expected grade")
```

The results are a lot messier for cultural differences, partly because of the opposite effect of foreign-born parents. In most cases, differences point to an in-between status when there is any daylight between the constituent groups. So, its the SES effects that are really pulling this toward the lower group.

## Summary table of resource differences

Lets see if I can put all of this into a simpler table that summarize the key issues.

```{r tab-resources}
temp <- coefs_cf %>%
  select(term, estimate) %>%
  mutate(estimate=round(estimate,3)*100,
         term=as.character(term)) %>%
  rename(race=term, held_back_cf=estimate)

tbl_resources <- acs_design %>%
  group_by(race) %>%
  summarize(fam_inc=round(survey_mean(family_income, vartype=NULL), 0),
            own_home=round(survey_mean(own_home, vartype=NULL), 3)*100,
            m_ba=round(survey_mean(degree_mother=="BA" | degree_mother=="G", 
                                   vartype=NULL), 3)*100,
            f_ba=round(survey_mean(degree_father=="BA" | degree_father=="G", 
                                   vartype=NULL), 3)*100,
            f_born=round(survey_mean(foreign_born, vartype=NULL), 3)*100,
            par_fb=round(survey_mean(foreign_born_father | foreign_born_mother, 
                                     vartype=NULL), 3)*100,
            spk_eng_well=round(survey_mean(spk_eng=="Yes",
                                           vartype=NULL), 3)*100,
            spk_eng_par=round(survey_mean(spk_eng_father=="Yes" & 
                                     spk_eng_mother=="Yes", vartype=NULL), 
                              3)*100) %>%
  mutate(race=as.character(race)) %>%
  left_join(temp) %>%
  arrange(held_back_cf)

index_biracial <- which(str_detect(tbl_resources$race, "/"))

tbl_resources %>%
  mutate_if(is.numeric, format, big.mark=",") %>%
  kbl(caption="Resource differences by race. Shading indicates biracial group. All results are survey weighted.",
      align = c("l",rep("r", ncol(tbl_resources)-1)),
      col.names=c("race","mean family income","% own home",
                  "% mother, BA degree","% father, BA degree",
                  "% foreign-born","% either parent foreign-born",
                  "% speak English well", "% Both parents speak english well",
                  "% held back, counterfactual")) %>%
  kable_paper() %>%
  row_spec(index_biracial, background = "#D3D3D3")

save(tbl_resources, file=here("analysis","output","tbl_resources.RData"))
```

The table sorts groups by their counterfactual probability of being held back. I think it does a nice job of summarizing the issues.

```{r time-end}
timestamp()
```

>>>>

---
title: "Main Analysis for Biracial Grade Retention Project"
output: 
  rmdformats::robobook:
    highlight: tango
    toc_depth: 2
    fig_height: 6
    fig_width: 9
    code_folding: hide
    lightbox: TRUE
---

@@

@@

@@
